### 2018-03-30 Paper - Latex/Paper - Skizze - v4.tex
Thesis - Result table ....

\input{arbeit-vorlage-praeambel.tex} % Importiere die Einstellungen aus der Präambel
% hier beginnt der eigentliche Inhalt

\author{Lydia Buntrock}
\title{master thesis}
\date{Januar 2018}

\hfuzz=\maxdimen \tolerance=10000 \hbadness=10000
% \usepackage[showframe=true]{geometry}

\begin{document}
  % Titelseite
  \begin{titlepage}
    \pagestyle{empty}
  	\begin{center}
      {\Large Freie Universität Berlin} \\
    	\begin{Huge}
      	Fachbereich Mathematik und Informatik \\
      	\vspace{3mm}
    	\end{Huge}
    	\vspace{2cm}
    	\begin{Large}
        \textbf{An analysis of maximum parsimony algorithms to predict parasitism in Eukaryota} \\
        \vspace{3mm}
        using a large multifurcated phylogenetic synthesis tree \\
    	\end{Large}
      \vspace{3cm}
      \textbf{Submitted on:} \\
      3 April 2018 \\
    	\vspace{2cm}
    	Lydia Buntrock \\
      E-Mail: info@irallia.de \\
     	\vspace{3cm}
      \textbf{Supervisor:} \\
      Prof. Dr. Bernhard Y. Renard \\
      \& \\
      Prof. Dr. rer. nat. Emanuel Heitlinger \\      
  	\end{center}
    \clearpage
    \pagenumbering{Roman}
  \end{titlepage}

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter*{Abstract}

  Parasitism can be defined as an interaction between species in which one of the interaction 
    partners, the parasite, lives in or on the other, the host. The parasite draws food from its 
    host and harms it in the process. According to estimates, above 40\% of all eukaryotes are 
    parasites. Nevertheless, it is computationally difficult to obtain information whether a 
    particular taxon is a parasite making it difficult to query large sets of taxa.

  Here we test in how far it is possible to use the Open Tree of Life (OTL), a synthesis of 
    phylogenetic trees on a backbone taxonomy (resulting in unresolved nodes), to expand available 
    information via phylogenetic trait prediction. We use the Global Biotic Interactions (GloBI) 
    database to categorise 25,962 and 34,860 species as parasites and free-living, respectively and 
    predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state information.

  We estimate the accuracy of our maximum parsimony based predictions using cross-validation and 
    simulation at 60-80\% overall, while strongly varying between clades. The cross-validation 
    results in an accuracy of 98.17\% which is explained by the fact that the data are not uniformly 
    distributed. We describe this variation across taxa as associated with available state and 
    topology information. We compare our results with several smaller scale studies, which used 
    manual expert curation and conclude that computationally inferred state changes largely agree in 
    number and placement with those. In clades in which available state information is biased 
    (mostly towards parasites, e.g. in Nematodes) phylogenetic prediction is bound to provide 
    results contradicting conventional wisdom.

  This represents, to our knowledge, the first comprehensive computational reconstruction of the 
    emergence of parasitism in eukaryotes. We argue that such an approach is necessary to allow 
    further incorporation of parasitism as an important trait in species interaction databases and 
    in individual studies on eukaryotes e.g. in the microbiome.

%---------------------------------------------------------------------------------------------------

  % This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
  %   Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  % The challenge here is the size of the tree and the little information about it. \\
  % Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
  %   tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  % For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
  %   \cite{Poelen2014} which we used, we could only collect 25,962 parasitic and 34,860 free-living 
  %   species. It follows that we have $\approx 2.4\%$ state information. \\
  % So far, especially small scale or highly manual studies have been carried out. In this scale, it 
  %   requires different data sources to be interconnected. \\
  % We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
  %   using the R package \textit{Castor} \cite{Louca2017}. \\
  % Nevertheless, the results are convincing and even though this is a purely computational approach 
  %   which did not include human experts input, results coincide with prior knowledge. Also regarding 
  %   the number of events, our estimates coincide with previous results by human experts, e.g. the 
  %   study by Weinstein and Kuris \cite{Weinstein2016}. \\

\tableofcontents
\clearpage
\pagenumbering{arabic}

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Introduction}
  This thesis is about the analysis of ancestral state reconstruction algorithms for non-binary 
    trees, applied to the currently largest phylogenetic synthesis tree of Open Tree of Life (OTL)
    \cite{Hinchliff2015} to predict parasitism in Eukaryota.

  For about 50 years, people have been working on ancestral state reconstruction, the inference of 
    evolution that leads to the given data. \\
  One of the first papers was written by Camin and Sokal, who in 1965 were working on algorithms for 
    discrete-state data \cite{Camin1965}. Different methods have been developed and the question is 
    which method is the most suitable for the problem at hand: the ancestral state reconstruction 
    for a huge non-binary tree with two discrete states.

% The first methods were only brute force \cite{Camin1965} 
%   Next came a set of parsimony algorithms such as: Fitch-parsimony \cite{Fitch1971}, 
%   Wagner-parsimony \cite{Swofford1987} ... \todo{weitere?}. \\
%   With more and more data, there is now the possibility to use more information to calculate the 
%   probabilities of the ancestral states. In addition to the states of the leafs, algorithms could 
%   also use branch lengths. The likelihood based algorithms came more in interest. \\

  Royer-Carenzi et al. distinguish two major classes of ancestral state reconstruction methods: \\
  The first is maximum parsimony: explain the current state with the least number of state changes 
    between the child and its ancestor. \\
  The other class they present describes modeling the character evolution as a stochastic process 
    and using the likelihoods to compute the possible ancestral character states. This is generally 
    done with a continuous time Markov model \cite{RoyerCarenzi2013}. \\
  \todo{Pasqualin et al. unterscheiden noch eine weitere Methode: stochastic mapping...}

  One of the major disadvantages of parsimony methods is that, unlike likelihood approaches, they 
    can not take divergence times (branch length) into account. Since the OTL does not include 
    development times of species, this can not be considered here. \\
  Another problem pointed out by Royer-Carenzi is that parsimony approaches are either based on 
    predefined parameters (generalized parsimony) or on strong and often controversial assumptions, 
    like irreversibility of transitions for dollo parsimony. This problem is unimportant to the 
    present task because in the analysis of the entire Eukaryota tree only generalized models are 
    meaningful. \\
  % Comparing the methods, we find that parsimony are methods in which you can not include information 
  %   such as branch length (development time of the species). This is also their main disadvantage 
  %   compared to the more complex likelihood based models. Since there are no branch lengths or other 
  %   information available for the present relationship tree, and we can thus not take advantage of 
  %   the other models. \\
  The comparison of methods gives us the maximum parsimony method as the simplest and fastest method 
    that meets all our requirements.
  % Parsimony-based methods are used in this work, since they are fast and fully sufficient for the 
  %   presented use case here, following the principle of the simpler model first. \\

  Felsenstein \cite{Felsenstein2003} discusses in his book two parsimony algorithms that generalize 
    previous methods (from Camin and Sokal \cite{Camin1965}, Farris \cite{Farris1970} and others): 
    Fitch parsimony \cite{Fitch1971} and Sankoff parsimony \cite{Sankoff1975}. Therefore, these are 
    the methods used in this work. \\
  % \anmerkungstext{Unter Farris war auch noch der Begriff Wagner trees in gebrauch, als 
  %   Verallgemeinerung der parsimonious trees von Camin und Sokal. (Lydia)} \\
  % \todo{Wagner-parsimony \cite{Swofford1987}} \\
   For Fitch, the algorithm has been extended from binary to non-binary trees. For the Sankoff 
    algorithm, Louca and Doebeli have presented an implementation for non-binary trees published in 
    an R package named \textit{castor} \cite{Louca2017}.

  For an ancestral reconstruction on the phylogeny of Eukaryota, we use Open Tree of Life (OTL) 
    \cite{Hinchliff2015}. \\
  Phylogeny describes the evolution of species, while the taxonomy is a classification according to 
    certain criteria in so-called taxa. \\
  OTL is a comprehensive, dynamic and digitally available tree of life constructed from published 
    phylogenetic trees along with a backbone of taxonomic data. Besides this, Hinchliff et al. also 
    offer an OpenTreeOfLife-Taxonomy (OTT) with the help of which we identify the individual nodes. \\
  It follows that the biggest 'phylogenetic tree' is this synthesis of phylogenetic trees filled 
    with a taxonomic trees given by OTL.

  For these large phylogenetic synthesis trees, however, ancestral state reconstruction has so far 
    only been done for Bacteria and Archaea for binary traits by Goberna and Verdú \cite{Goberna2015}.
    However, this differs from eukaryotes in the sense that complex traits such as parasitism depend
    on more than one gene.

  The present tree structure of OTL is not binary but multifurcated, meaning that each node has
    multiple ($n > 2$) children or in other words its degree, number of adjacent nodes, is greater 
    than 3 \cite{Felsenstein2003}. \\
  Parsimonious in phylogeny refers to favoring the tree that needs the least evolutionary change to 
    explain the observed data. Maximum parsimony methods have been developed for phylogenies, which 
    are usually depicted as binary trees. Therefore, the selected parsimony methods are not directly 
    applicable, as they were specifically applied to much smaller subtrees, where all splits are 
    known. \\
  We will extend and test the existing maximum parsimony algorithms of Fitch \cite{Fitch1971} and 
    Sankoff \cite{Sankoff1975} for this task and estimate their predictive power.

  The Sankoff method is implemented by Louca et al. for the non-binary case and is available as an R 
    package called \textit{Castor} \cite{Louca2017}. \\
  Note, the original Fitch algorithm has the sole purpose of minimizing the number of transitions 
    and not reconstructing the ancestral nodes. Felsenstein \cite{Felsenstein2003} describes a 
    simple extension for the reconstruction. In this work, the algorithm extended to reconstruction 
    is adapted to multifurcated trees, based on the critical reevaluation of this extension by 
    Cunningham et al. \cite{Cunningham1998}.  
  % In addition, we have implemented the Fitch 
  %   method \cite{Fitch1971} \todo{hier vll die andere quelle nutzen nach der wir uns gerichtet haben?}and adapted it for multifurcated trees. \\% to make it usable and comparable for our current problem.
  % This achieves that we can predict ancestral states and unknown states of living species for large 
  %   non-binary trees.

  To accomplish this task, next to the phylogenetic tree information about the current species' 
    states is needed. \\
  Most of the largest interaction databases are offline or outdated. For example: IWDB (Interaction 
    Web Database) \cite{IWDB2003}, Webs on the Web \cite{WOW2004}, Animal Diversity Web 
    \cite{Myers2003} and ecoweb \cite{Cohen2010}. \\
  We use the interaction database Global Biotic Interactions (GloBI) \cite{Poelen2014} because it is 
    including most of the known databases and is still growing actively \cite{Poelen2014}. \\
  The data in GloBI are stored as interactions e.g.  species A parasitize species B. From this we 
    conclude that species A is parasitic and species B free-living. \\
  From this data, we could specify $\sim$2.3 million leaf nodes 34,860 as free-living and 25,962 
    as parasitic.

  There are many different ways to define parasitism. Since we use GloBI to classify species, we use 
    their definition of parasitism. Again, in GloBi, Ontobee definitions are used \cite{Xiang2011}. \\
  The interaction \textit{has parasite} is defined as: "An interaction relationship between two 
    organisms living together in more or less intimate association in a relationship in which 
    association is disadvantageous or destructive to one of the organisms."\footnote{
      \hyperlink{http://www.ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO_0002445}
      {ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO\_000244}; Last checked: 22.03.2018.
    }. This definition includes: ecto- and endoparasites, parasitoids, kleptoparasites and pathogenes. \\

  The objectives of this work are the following points: (1) Find a suitable ancestral state 
    reconstruction method. (2) Accomplish reconstructing on the Eucaryotic synthesis tree of OTL. \\
  The goal of Point 1 is to evaluate the possible methods based on a simulation of our data 
    situation. \\
  The Sankoff algorithm implemented by Louca et al. is the best in our comparisons. Therefore, 
    point 2 consists of reconstructing the ancestral states and predicting the unknown leaf states 
    with the aid of this algorithm. And then perform an evaluation of the results. \\

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Aims}
  The aim of this thesis is the application of maximum parsimony algorithms to non-binary trees and 
    very large data sets to predict unknown leaf node states. \\
  We discuss different ancestral state reconstruction methods and test the most appropriate ones on 
    our data. Hereby, we want to find out how far it is possible to use the Open Tree of Life (OTL) 
    to expandavailable information via phylogenetic trait prediction. As a basis for known states, 
    we use the Global Biotic Interactions (GloBI) database, which cathegorises 34,860 and 25,962 
    species as free-living and parasites. This data is used to perform an ancestral state 
    reconstruction and predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state 
    information.

  In order to generate a realistic simulation, influencing parameters are investigated. \\
  Since the transitions are minimized in an ancestral state reconstruction, this is an important 
    parameter to consider. On the other hand, the completeness of our input data is an influencing 
    value. Therefore, two major types are distinguished:
  \begin{enumerate}
    \item Biological parameters (a result of the evolutionary process):
      \begin{itemize}
        \item State distributions and transition probabilities
      \end{itemize}
    \item Distribution of missing information:
      \begin{itemize}
        \item Lack of information on topology ($\rightarrow$ multifurcations)
        \item Lack of information of states of some leaf nodes
      \end{itemize}
  \end{enumerate}

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Methods}
  In this thesis, a maximum parsimony algorithm is applied to the Eukaryota tree to obtain an 
    ancestral state reconstruction of free-living versus parasite states. \\
  This chapter is divided into the following sections: the description of the data sources 
    (section \ref{sec:methods - data sources}) used and the analysis of these data as a preparation 
    for the simulation (section \ref{sec:methods - metadata analysis}). A description of the 
    analyzed methods for the ancestral state reconstruction (section 
    \ref{sec:methods - ancestral state reconstruction}) and then an explanation of how they fit the 
    problem at hand (section \ref{sec:methods - simulation}). And at the end, the real data analysis 
    with the Sankoff method (section \ref{sec:methods - real data analysis}). \\
  Figure \ref{fig:workflow} briefly outlines these relationships. A more detailed view of the 
    workflow can be found in the appendix \ref{fig:BigWorkflow}.
  \begin{figure}[h!]
    \centering
    \includegraphics[width=0.6\textwidth]{Figures/Workflow-overview.png}
    \caption{The Workflow of the resulting procedure with the following steps: \\
      (1) Retrieve phylogenentic tree data as input for the tree (OTL) and the state information (GloBI).
      (2) Get metadata of these for a realisitc simulation of the maximum parsimony algorithms (Fitch \& Sankoff).
      (3) Build and run the simulation.
      (4) Evaluation of parameters for the simulation and the ancestral state reconstruction of the real tree.
      (5) Evaluate the accuracy of developed algorithms and choose the best.
      (6) Run the resulting algorithm on the original data.
      (7) Evalute and interprete the results. %$\rightarrow$ Origins etc...
    }
    \label{fig:workflow}
  \end{figure}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Data sources} \label{sec:methods - data sources}
    Two types of data are needed for an ancestral state reconstruction: a tree and information about 
      the states.

    For the subsequent analyzes on the Eukaryote tree the database of Open Tree of Life (OTL) is used 
      (downloaded on 16.02.2018) \cite{Hinchliff2015}. This database gives a synthesis of phylogenetic 
      trees (currently 819 trees) and a taxonomic tree\footnote{
        \hyperlink{https://tree.opentreeoflife.org/about/synthesis-release/v9.1}
        {https://tree.opentreeoflife.org/about/synthesis-release/v9.1}; Last checked: 22.03.2018.
      }. OTL also  includes the large phylogenetic database TreeBASE \cite{Hinchliff2015}.

    Furthermore the Open Tree Taxonomy (OTT) from OTL is used because it includes most of the known 
      taxonomies and is synthesised by prefering taxonomies that match with available phylogenetic 
      data. The team from OTL prefer a maximum number of species \cite{Hinchliff2015}, this results 
      in the synthesis between taxonomy and phylogeny.

    For the state information the Global Biotic Interactions database (GloBI) is taken 
      \cite{Poelen2014} (downloaded on 29.01.2018). \\
    This database consists of entries of the form: species A (source) interacts with B (target). \\
    A number of interactions have been identified\footnote{\hyperlink{
      https://github.com/jhpoelen/eol-globi-data/blob/master/eol-globi-lib/src/main/java/org/eol/globi/domain/InteractType.java
      }{https://github.com/jhpoelen/eol-globi-data/.../InteractType.java}; Last checked: 22.03.2018.}
      , including those indicating whether the species source or target has become a parasite or a 
      free-living species from the biological perspective. They are the following:
    \begin{itemize}
      \item free-living source: preysOn, eats, flowersVisitedBy, hasPathogen, pollinatedBy, 
        hasParasite, hostOf
      \item free-living target: preyedUponBy, parasiteOf, visitsFlowersOf, pathogenOf, hasHost
      \item parasite source: parasiteOf, pathogenOf
      \item parasite target: hasParasite, hasPathogen
    \end{itemize}
    Of these interactions, e.g. species A parasitize species B, the state of the species is 
      determined, here is species A parasitic and species B free-living. In the case that a parasite 
      conquers (parasitizes) another parasite yields conflicting states for the second species. 
      This is solved by preferring parasitic. \\
    For each species known identifyers are stored in GloBI. This includes OTT (the taxonomy of OTL). 
      All species that have stored an OTT indentifyer and have a matching interaction are divided into 
      two lists: parasites and free-livings.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Metadata analysis} \label{sec:methods - metadata analysis}
    Based on the input data, generalized linear models are compared with poisson respectively binomial 
      regression according to their residuals. In order to compare models of different complexity, the 
      BIC (Bayesian Information Criterion) values are calculated in addition to the residuals. \\
    There are two different information criteria: AIC (Akaike Information Criteria) and BIC. \\
    The advantage of the BIC is that the penalty is dependent on the sample size and is therefore 
      advantageous for large samples. \\
    For all these calculations, the following R functions are  used: \textit{glm()}, \textit{anova()} 
      and \textit{BIC()}.
    % https://stat.ethz.ch/R-manual/R-devel/library/stats/html/glm.html

    For each node, depth, min, max and mean height are  noted. The node depth is calculated as the
      distance (number of edges) to the root node. The node height is calculated in three different 
      ways: min, max and mean height refers to the (mininmal, maximal, mean) distance to a leaf node. \\
    The influence in the modeling of these parameters is tested, additively as well as 
      multiplicatively to the models.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Ancestral state reconstruction methods} \label{sec:methods - ancestral state reconstruction}
    The methods used to reconstruct the ancestral states are Fitch parsimony \cite{Fitch1971} and 
      Sankoff parsimony \cite{Sankoff1975}, which are explained in the following subsections.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Fitch maximum parsimony}
      Based on the work of Cunningham et al. \cite{Cunningham1998} the Fitch method \cite{Fitch1971} 
        is implemented and extended by us for the multifurcated tree case. To understand the 
        differences to the multifurcated case, the algorithm for the binary case is briefly explained 
        and referred to the extension.

      Input: A rooted, binary tree, with state informations in the leaf nodes. Each node is depicted as 
        a set of states. There are only two states in this thesis, free-living (FL) and parasitic (P). 
        Internal nodes have three sets, which are empty at the beginning, excluding the root node, it 
        has only one. Leaf nodes have their state as a set (e.g. \{FL\} or \{P\}, unknown leaf nodes the 
        union of all possible states (\{FL, P\}).

      The algorithm traverses three times through the tree and fills these sets. \\
      In each step, two sets are considered and their intersection is formed. There are two cases:
      \begin{enumerate}
        \item The intersection is not empty and corresponds to the new set.
        \item The intersection is empty. $\rightarrow$ Build the union of these sets as new set.
      \end{enumerate}
      First traverse from the leaf nodes to the root: each internal node is formed from its child nodes, 
        where the only information lies at the beginning. \\
      Second traverse from the root node to the leafs: each internal node is formed from its parent node 
        and its sibling node. \\
      Last traversion (direction does not matter): build the final state for every node. It is formed 
        from the sets of previous traversals.

      (The original Fitch algorithm is designed to minimize transitions without predicting actual states 
        of internal nodes, so it is just the first traversal.) \\
        \begin{wrapfigure}[20]{r}{0.5\textwidth}
          \centering
          \includegraphics[width=0.4\textwidth]{Figures/Fitch1.png}
          \caption{Fitch algorithm for binary trees. \\
            The unknown leaf node is discribed with both states. Computed internal nodes (exclusive the 
            root node) consists of three sets, where the last set is the final one (bold). \\
            From the second internal node (seen from the root node) there are several possibilities to 
            create the second and third set.}
          \label{fig: binary Fitch}
        \end{wrapfigure}
      The extension to the non-binary case is quite obvious, but holds some opportunities. In this case, 
        more than two children may be present for the first traversal, but the incision or union may 
        also be formed over more than two sets. Also in the second traversing, there may be several 
        sibling nodes. However, there are several possibilities here that are all tested and compared 
        in the simulation. Some of these options are already available in the binary case:
      \begin{itemize}
        \item The parent node has (except for the root node) two state sets, because he came through 
          the up-traversing previously. Are both sets used or only the first traversing?
        \item If there are several siblings, is the cut or union made first of these, or is everything 
          made directly including the parent node.
      \end{itemize}
      The first point already has an effect on the binary case. Figure \ref{fig: binary Fitch} shows 
        both possibilities of the three sets. \\
      Cunningham uses only the first state set of the parent node \cite{Cunningham1998}. \\
      From these two points four different versions of Fitch are formed:
      \begin{enumerate}
        \item Fitch 1: First state set of parent node; intersection/union of siblings first.
        \item Fitch 2: First state set of parent node; intersection/union of siblings together with parent node.
        \item Fitch 3: Both state sets of parent node; intersection/union of siblings first.
        \item Fitch 4: Both state sets of parent node; intersection/union of siblings together with parent node sets.
      \end{enumerate}

      \begin{figure}[h!]
        \centering
        \includegraphics[trim = 24mm 0mm 28mm 14mm, clip, width=\textwidth]{Figures/simulation_fitch_evaluation.png}
        \caption{Test of Fitch Versions. \todo{beschriften!!}}
        \label{fig:Fitch versions}
      \end{figure}
      These four versions are tested in the simulation with 100 trees and 10000 leaf nodes and
        distributions of 70\% FL to 30\% P and 60\% FL to 40\% P. Figure \ref{fig:Fitch versions} shows 
        this over all unknown node percentage. \\
      % At 90\% unknown nodes and 90\% of multifurcation of the internal nodes, version 1 is 89.26\%, 
      %   version 2 is 89.26\%, version 3 is 89.35\%, and version 4 is 89.31\% correct. Therefore, 
      %   only version 3 is used for all further simulations.
      At 95\% unknown nodes and 95\% of multifurcation of the internal nodes, version 1 is 88.37\%, 
        version 2 is 88.37\%, version 3 is 88.4\%, and version 4 is 88.39\% correct. Therefore, 
        only version 3 is used for all further simulations.

    % Fitch versions:
    % | 88.37 % | 88.37 % | 88.4 % |88.39 % |

    %---------------------------------------------------------------------------------------------------
    \subsubsection{Sankoff}
      Maximum parsimony algorithm from Sankoff implemented in the R package castor \cite{Louca2017}.
        From this the function \textit{$hsp\_max\_parsimony()$} is used with default settings including
        $transition\_costs="all\_equal"$. \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Simulation}\label{sec:methods - simulation}
    The simulation compares these different ancestral state reconstruction algorithms with each 
      other. \\
    First different implementations of the Fitch maximum parsimony are compared and then the best of 
      them is compared with the implementation of the Sankoff algorithm of the \textit{Castor} package 
      \cite{Louca2017}.

    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.95\textwidth]{Figures/Workflow-Simulation.png}
      \caption{A simulation is performed to compare different maximum parsimony algorithms. \\
        The course of the simulation with influence of the metadata analysis from the real data can
          be seen: \\
        (1) A phylogenetic tree is created randomly. (2) Simulate node states for all nodes. (3) 
          'Forget' internal states and some leaf node states. (4) 'Lose' phylogeny information. (5) 
          Make a nodelist for the algorithm. (6) Run algorithms. (7) Evaluate results. \\
        Points 3 and 4 are influenced by metadata of the real-data analysis.}
      \label{fig:Simulation Workflow}
    \end{figure}
    The course of a simulation is shown in Figure \ref{fig:Simulation Workflow}. The individual steps 
      are explained in the following.

    A tree is needed to perform a simulation of ancestral state reconstruction. It has to be decided 
      whether to take the real tree or simulate a tree. In this simulation, trees are created randomly, 
      as one can replicate a complete binary phylogentic tree. Thus, there is also the possibility to 
      simulate the multifurcation. \\
    To get a random binary tree, the Phylo package from biopython is used \cite{Cock2009}. They offer 
      a \textit{randomized()} function which returns a BaseTree\footnote{
        \hyperlink{https://github.com/biopython/biopython/blob/master/Bio/Phylo/BaseTree.py}
        {https://github.com/biopython/biopython/blob/master/Bio/Phylo/BaseTree.py}
      }.

    The next step is to simulate states for all nodes. \\
    Here the influence of the biological parameters as transition probabilities and distributions of
      \begin{wrapfigure}[21]{r}{0.55\textwidth}
        \centering
        \includegraphics[width=0.55\textwidth]{Figures/40-60_all.jpg}
        \caption{Set node states: Distribution of states (left); traversion through the tree (right). \\
          Start with a free-living root node (FL: red). \\
          (1) + (2) Draw random numbers for its children from the free-living distribution (red), the numbers are under the threshold 
          $\rightarrow$ the nodes are again free-living; go on with the children. \\
          (3) The number drawn is above the threshold. $\rightarrow$ The node state changes to parasitic (P: blue). \\
          (4) Now draw random numbers from the parasite distribution (red) until one number lies unter the threshold. Then change back.}
        \label{fig:set node states}
      \end{wrapfigure} 
      states is included. \\
    Since there is no statement about general transition probabilities, these are all set the same: \\
    $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$.

    For the distributions of states different ratios of parasites (P) to free-livings (FL) are 
      simulated with the help of beta distributions and a given threshold:
      \begin{itemize}
        \item 50\% P to 50\% FL,
        \item 40\% P to 60\% FL,
        \item 30\% P to 70\% FL and 
        \item 20\% P to 80\% FL.
      \end{itemize}

    Procedure: The root node is defined as ancestor of all subsequent species and in this case, 
      determined to be free-living. Therefore, the beta distribution for free-living is used at the 
      beginning. Now traverse from the root to the leaf nodes, always pulling out of the current 
      distribution until the randomly drawn number is above the threshold and the new node changes 
      state. Figure \ref{fig:set node states} shows a part of this simulating states and the 
      associated distributions.
    % To ensure that the parameter of the binomial distribution is restricted to the [0,1] interval, it 
    %   is modeled with a beta distribution as in Figure \ref{fig:set node states}. \\

    After traversing through the tree, each state is saved in a nodelist associated with the node 
      indentifyer which is the OTT from OTL.

    Here begins the simulation of the lack of information, such as the lack of information on topology 
      (multifurcation) and of states of some leaf nodes.
    % = In the simulation, the influence of the multifurcation and missing data in leaf nodes on the 
    %   predictive accuracy of the ancestral state reconstruction algorithms is tested.

    In the real tree, there is usually only information about species living today $\rightarrow$ leaf 
      nodes. And beyond only a small percentage of these. All information about the states of the 
      internal node and one leaf node is removed and stored in another column to the node.

    Last step for the preparation is the multifurcation of the tree. As previously explained, some 
      divisions in the tree are not known, so the real tree is not binary. This multifurcation is 
      simulated by an equally distributed percentage of forgotten internal nodes.

    Different percentages of removed information are simulated.

    The last step is the evaluation of the results. This is done with a simple difference calculation 
      of the node states. \\
    In the nodelist, the originally simulated states and the newly calculated states are stored for 
      each node ($FL = 0$, $P = 1$). The sum of the differences of the node states gives the distance 
      of the prediction to the original tree.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Real data analysis} \label{sec:methods - real data analysis}
    For the evaluation of real data results, a leave-100-out cross-validation is performed. For this 
      purpose 100 randomly distributed 100 states are left out.

    Analogous to the missing leaf node states, generalized linear models with binomial regression are 
      compared according to their residuals, based on the "true" / "false" prediction. \\
    Again, in order to compare models of different complexity, the BIC (Bayesian Information Criterion) 
      values are calculated in addition to the residuals. \\
    For all these calculations, the following R functions are used: \textit{glm()}, \textit{anova()} 
      and \textit{BIC()}. \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Implementation}
    The complete code is located on GitHub: 
      \hyperlink{github.com/Irallia/IZW-HU-Parasites}{github.com/Irallia/IZW-HU-Parasites}. \\
    Most of the code is written in Python, the analyzes and the use of the Castor package in R. Some 
      shell scripts are used to execute whole workflows.

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Results}
  This work deals with the ancestral state reconstruction of the entire Eukaryota relatives tree.

  For this reconstruction, we first analyzed our data. This is the tree of OTL \cite{Hinchliff2015}
    and the data for the leaf node states (free-living or parasite states) from GloBI 
    \cite{Poelen2014} (section \ref{sec:results - missing information}), where we have determined 
    that the entire Eukaryota tree is 89.45\% multifurcated (missing internal nodes) and that we do 
    not have any state information for 97.34\% nodes. \\
  In addition, we tested different models for these two paramters and found out that multifurcation 
    can best be modeled with the taxa 'order' multiplied by the mean height of the knots. Also, the 
    missing states could be modeled with the taxa 'order' and in this case multiplied by the depth 
    of the nodes.

  Next we compared different possible methods. We decided to take a closer look at maximum parsimony 
    algorithms because they are fast and best suited to the problem at hand. We tested these on 
    different simulations of the data and compared their predictive power (section 
    \ref{sec:results - simulation}). \\
  We compared them with the result that the Sankoff algorithm performs significantly better than 
    Fitch. The multifurcation rate has little impact on Sankoff, but Fitch's prediction 
    significantly breaks down at more than 60\% missing internal nodes. However, the amount of 
    missing node state information has a linear impact on both, as does the ratio of free-living to 
    parasitic (the more balanced, the worse is the prediction).

  As a result, we used the Sankoff algorithm \cite{Sankoff1975} implemented by Louca et al. 
    \cite{Louca2017}, which performed best, for the actual reconstruction on the real data. \\
    \todo{... biological view}
  %  On the one hand, we have examined the results of some 
    % subtrees for their credibility with the help of biological background knowledge (subsection 
    % \ref{sec:results - biological view}). On the other hand, we
  We evaluated these by performing a leave-100-out cross-validation and predicted approximately 
    98.17\% of the omitted states correct (subsection \ref{sec:results - cross-validation}). \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Availability of internal nodes and state information} \label{sec:results - missing information}
    As previously presented, we have two types of missing information: unknown states of leaf nodes 
      and multifurcation.

    A tree is multifurcated if there are nodes that have more than two children. A binary tree with 
      $n$ leaf nodes has $n-1$ internal nodes. The present Eukaryota tree of OTL has 2,293,463 leaf 
      nodes and only 241,974 internal nodes, that is:
    $$100-\frac{100}{(2293463-1) \times 241974} \approx 89.45\%$$
      missing internal nodes. This means that there is a lack of information about the underlying 
      phylogeny. Instead of a binary tree this tree is highly multifurcated.

    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l||r|r| }
          \hline
          \bfseries Subtree of & \bfseries Unknown States & \bfseries Multifurcation \\ 
          \hline \hline
          Eukaryota       & 97.34\%  & 89.45\% \\
          \hline \hline
          Metazoa         & 96.44\%  & 87.93\% \\ \hline
          Fungi           & 98.87\%  & 96.97\% \\ \hline
          Chloroplastida  & 99.14\%  & 89.46\% \\
          \hline \hline            
          % Arthropoda      & 97.49\%  & 89.95\% \\ \hline
          Apicomplexa     & 86.26\%  & 87.16\% \\ \hline
          Nematoda        & 89.01\%  & 88.59\% \\ \hline
          Chordata        & 88.59\%  & \cellcolor{green!50}66.49\% \\ \hline
          Platyhelminthes & \cellcolor{green!50}68.73\%  & 80.34\% \\
          \hline \hline            
          Insecta         & 97.11\%  & 90.78\% \\
          \hline  
        \end{tabular}
      \end{center}
      \caption{Examination of subtrees regarding missing information. \\
        The percentage values show the proportion of missing information of: unknown states (missing state 
          information of leaf nodes) and multifurcation (missing internal nodes). \\
        The subtrees are from different taxa: domain (Eukaryota), kingdom (Metazoa, Fungi, 
          Chloroplastida), phylum (Apicomplexa, Nematoda, Chordata, Platyhelminthes) and class 
          (Insecta). \\
        The two by far smallest values are highlighted in green.} 
      \label{table:percentage loss information subtrees} 
    \end{table}

    For the missing state information the present Eukaryota tree with 2,293,463 leaf nodes, 34,869 
      free-livings and 25,962 parasites are found. This gives
      $$100-\frac{100}{2293463 \times (34860+25962)} \approx 97.34\%$$
      unknown states of leaf nodes.

    We calculated these percentages of missing nodes and also missing state information for some 
      subtrees and plotted them in table \ref{table:percentage loss information subtrees}.

    \begin{figure}[h!]
      \centering
      \begin{subfigure}[b]{0.59\textwidth}
        \includegraphics[width=0.9\textwidth]{Figures/multifurc.pdf}
        \caption{Histogram with automatic binwidth. \\ ~}
      \end{subfigure}
      \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[trim = 0mm 0mm 30mm 0mm, clip, width=0.9\textwidth]{Figures/multifurc_small.pdf}
        \caption{Histogram with $binwidth = 1$. Higher multifurcation than 30 has been cut off. \\
          light blue: binary; \\ dark blue: multifurcation}
      \end{subfigure}
      \caption{Histograms showing the multifurcation of the internal nodes of the synthesis tree. \\
        For each node, the number of children (degree $-1$) is collected. A node is multifurcated 
        if it has more than two children, so we deducted two from each number of children. The two 
        histograms show: number of children $-2$ on the x-axis with log scale and the number of nodes 
        with this amount on the y-axis.}
      \label{fig:childrenOfNodes}
    \end{figure}

    For a first overview, we collected for each node its number of children (degree $-1$), 
      and plotted this in two histograms, see figure \ref{fig:childrenOfNodes}.

    The multifurcation affects only the internal nodes. We collected the number of children $-2$ of 
      these nodes (because a node with two children is binary). That means it describes the number of 
      nodes which we have lost from the real (binary) phylogenetic tree. \\
    It can be recognized that we are very far from a binary tree.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Comparison of different models for multifurcation}
      For the regression analysis of the multifurcation we set up several generalized linear models that 
        could describe multifurcation. \\
      In doing so, we allowed the different influence of the taxa and the heights and depths of a node 
        to be included. From this we got 9 times 4 models of different complexity levels (first row) and 
        the associated BIC values (table \ref{table:BIC multifurcation}).

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & 8273333 & \cellcolor{green!15}7937828 & \cellcolor{green!20}7842157 & \cellcolor{green!30}7644249 \\
            multifurc $\sim$ taxa & 8257680 & \cellcolor{green!15}7922207 & \cellcolor{green!20}7826490 & \cellcolor{green!35}7574154 \\
            \hline
            multifurc $\sim$ taxa + depth & 8273318 & \cellcolor{green!15}7934322 & \cellcolor{green!20}7839364 & \cellcolor{green!35}7539999 \\
            multifurc $\sim$ taxa + max.height & \cellcolor{green!15}7993515 & \cellcolor{green!25}7749121 & \cellcolor{green!30}7661817 & \cellcolor{green!40}7416211 \\
            multifurc $\sim$ taxa + min.height & 8251211 & \cellcolor{green!20}7875521  & \cellcolor{green!25}7778327 & \cellcolor{green!35}7516883 \\
            multifurc $\sim$ taxa + mean.height & \cellcolor{green!20}7825417 & \cellcolor{green!30}7644249 & \cellcolor{green!35}7572474 & \cellcolor{green!45}7340741 \\
            \hline
            multifurc $\sim$ taxa * depth & 8235932 & \cellcolor{green!20}7836755 & \cellcolor{green!25}7757688 & \cellcolor{green!45}7383808 \\
            multifurc $\sim$ taxa * max.height & \cellcolor{green!15}7963438 & \cellcolor{green!30}7693555 & \cellcolor{green!30}7614820 & \cellcolor{green!45}7335338 \\
            multifurc $\sim$ taxa * min.height & 8214030 & \cellcolor{green!20}7808940 & \cellcolor{green!30}7690618 & \cellcolor{green!45}7336627\\
            multifurc $\sim$ taxa * mean.height & \cellcolor{green!25}7768360 & \cellcolor{green!35}7536296 & \cellcolor{green!50}7484953 & \cellcolor{green!50}7206369 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of the multifurcation models. \\
          These models are created with the R function \textit{glm()} and compared with the \textit{BIC()} 
            function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC multifurcation} 
      \end{table}
      Within every complexity class it can be seen that the mean height gives the best additional factor. \\
      Despite higher complexity, the BIC values are getting smaller from model to model, meaning that 
        the finest model available here is also the best one of these. Lower taxa than orders (e.g. 
        family) are computationally too expensive to calculate. \\
      The model \textit{multifurc $\sim$ order * mean.height} turns out to be the best of our models, 
        whereby it is possible that e.g. \textit{multifurc $\sim$ family * mean.height} is better. \\

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Comparison of different models for missing state information}
      Next to the problem of the multifurcation of the tree is the little interaction data thet we have 
        for the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 free-living species and 47,332 parasite species 
        for the whole tree of life. From the Eukaryota we could determine 25,962 and 34,860 species as 
        parasites and free-living. With 2,293,463 leaf nodes we still have about 97.34\% unknown leaf 
        nodes.

      We also compared different models in terms of their BICs (Table: \ref{table:BIC unknown information}). \\ 
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |l|r|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class & \bfseries Order \\
            \hline \hline
            multifurc $\sim$ taxa & \cellcolor{green!15}545799 & \cellcolor{green!35}500004 & \cellcolor{green!45}485121 & \cellcolor{green!45}484681 \\
            \hline
            multifurc $\sim$ taxa + depth & \cellcolor{green!15}544862 & \cellcolor{green!40}493808 & \cellcolor{green!45}481869 & \cellcolor{green!50}478851 \\
            \hline
            multifurc $\sim$ taxa * depth & \cellcolor{green!15}544179 & \cellcolor{green!45}489845 & \cellcolor{green!45}481494 & \cellcolor{green!50}478188 \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values unknown state information models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC unknown information} 
      \end{table}

      It also follows from this table that the most complex model is the best. In general, the BIC 
        values are smaller than those of the multifurcation models. The modeling here is thus better.
      Again, the calculation of finer models (e.g. family) was too expensive.

      These missing data modeling results could be used to better simulate the data. \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Influence of different parameters on the prediction} \label{sec:results - simulation}
    As presented, we compare two methods in our simulation to their prediction accuracy: Fitch and 
      Sankoff.

    First, we tested different distributions of parasites to free-livings including threshold (first 
      column). It can be observed that the more balanced the percentages of free-livings and parasites, 
      the worse the prediction of the algorithms (second and third column). \\
    It can be seen that the predictive power of Sankoff is always greater or equal to the percentage
      of free-living distribution, and therefore more accurate than guessing. \\
    On the other hand, we examined the influence of missing internal nodes (ridge of multifurcation) 
      and missing leaf node information (unknown leaf nodes). \\
    Both factors have a relatively linear influence on the Sankoff method. Fitch, on the other hand, 
      breaks significantly in his prediction from about 70\% unknown leaf nodes or mutifurcation. \\
    Figure \ref{fig:influence of unknown data} shows the results of examining various parameters.

    Since we have a lot of missing data in most subtrees (both internal nodes and state informations), 
      the Sankoff gives a better prediction and is thus used for real data analysis.

    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 100mm 0mm 0mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
    \end{figure}
    ~ \\
    \begin{figure}[h!]
      \centering
      \includegraphics[trim = 0mm 0mm 0mm 150mm, clip, width=\textwidth]{Figures/simulation_evaluation_1.png}
      \caption{Influence of unknown data to prediction. \\
        The first column describes the distributions of free-livings and parasites with a given 
          threshold for the respective simulations to the right. Note, these distributions are chosen 
          to form the specified proportions with the threshold, ignoring the number of state changes. \\
        The middle column investigates the influence of the unknown states, the right the influence of
          the strength of the multifurcation. \\
        The y-axes indicate the percentage of correctly predicted states (including known states). On 
          the x-axis the percentage of forgotten states or missing internal nodes. \\
        Each point corresponds to the average of one hundred simulations, each with 10,000 leaf nodes. \\
        For the middle column we set the strength of the multifurcations to 0.95\%, similar to the 
          real data and in the right column the amount of the unknowns to 0.95\% also similar to the 
          real data.}
      \label{fig:influence of unknown data}
    \end{figure}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Results of the real data analysis created with Sankoff}
    This section is about evaluating the prediction of real data using the Sankoff method. \\
    It is divided into two subsections. \ref{sec:results - biological view}: The analysis of some subtrees using 
      biological background knowledge. \ref{sec:results - cross-validation}: Presentation of the results 
      of the leave-100-out cross-validation.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Biological view} \label{sec:results - biological view}
      To analyze the results, we have selected some phyla (subtrees) to evaluate our results 
        selectively from the biological point of view: Chordata, Nematoda, Platyhelminthes and 
        Apicomplexa. \\
      In Table \ref{table:phylum leaf nodes states} we compare the given states with the predicted ones. \\
      Several factors such as the distribution of existing input data via parasitics and free-livings, 
        faulty input data from GloBI and reinforcement of errors by multifurcation play an important 
        role, which crystallize out in these examples. \\
      An important factor here is the credibility of the results. The accuracy of the input data 
        stands and falls with the presence and the correctness of the data of GloBI. Errors of incorrect 
        input data can be amplified by incorrect prediction of unknown species and can be reversed in 
        order to improve the data situation of GloBI. \\
      Since we look at such large trees we can not expect to know all the parasites, so we look at 
        individual positives. This is positive in the sense that the majority have the opposite state.

      \newcommand*\rot{\rotatebox{90}}

      % \tiny	\scriptsize \footnotesize \small \normalsize \large \Large \LARGE \huge \Huge
      \begin{table}
        \begin{center}
          \hspace*{-0.5cm}\begin{tabular}{|>{\scriptsize}c|>{\scriptsize}l|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r||>{\scriptsize}r|>{\scriptsize}r|>{\scriptsize}r| }
            \hline
            & \bfseries Taxa & \bfseries \#internal & \bfseries \#leaf 
                & \multicolumn{2}{>{\scriptsize}c||}{\bfseries known states} & \multicolumn{3}{>{\scriptsize}c|}{\bfseries final states}
                & \bfseries Root & \bfseries \#origins & \bfseries \#losses \\
            & \bfseries subtree & \multicolumn{2}{>{\scriptsize}c||}{\bfseries nodes} & \bfseries FL & \bfseries P
                & \bfseries [0, 0.5) & \bfseries 0.5 & \bfseries (0.5, 1]
                & \bfseries node & \multicolumn{2}{>{\scriptsize}c|}{\bfseries (with rounding)} \\
            & \bfseries & & & & 
                & \bfseries FL & & \bfseries P 
                & state & \bfseries FL $\rightarrow$ P & \bfseries P $\rightarrow$ FL \\
            \hline \hline
            \parbox[c][7mm][c]{1mm}{\multirow{2}{*}{\rot{\bfseries Domain~}}}
            & Eukaryota & 241974 & 2293463      & 34860 & 25962     & 2053212 & 7775 & 474450     & 1 & 462 & 369 \\
            & & &                               & 57.31\% & 42.69\% & 80.98\% & 0.31\% & 18.71\%  & P & & \\
            \hline \hline
            \parbox[c]{1mm}{\multirow{6}{*}{\rot{\bfseries Kingdom}}}
            & Chloro- & 43486 & 416478          & 3519 & 77         & 410795 & 4182 & 1501        & 0.5 & 97 & 222 \\
            & plastida & &                      & 97.86\% & 2.14\%  & 98.63\% & 1.00\% & 0.36\%   & & & \\ \cline{2-12}
            & Fungi & 9534 & 31457              & 577 & 2983        & 38520 & 5723 & 266463       & 0 & 42 & 2 \\
            & & &                               & 16.21\% & 83.79\% & 12.40\% & 1.84\% & 85.76\%  & FL & & \\ \cline{2-12}
            & Metazoa & 179944 & 1491012        & 30758 & 22373     & 1329065 & 25535 & 136412    & 0.5 & 321 & 129 \\
            & & &                               & 57.89\% & 42.11\% & 89.14\% & 1.71\% & 9.15\%   & & & \\
            \hline \hline
            \parbox[c]{1mm}{\multirow{10}{*}{\rot{\bfseries Phylum}}}
            & Apicomplexa & 239 & 1863          & 1 & 255           & 1 & 0 & 1862                & 1 & 0 & 1 \\
            & & &                               & 0.39\% & 99.61\%  & 0.05\% & 0\% & 99.95\%      & P & & \\ \cline{2-12}
            & Arthropoda & 120479 & 1198981     & 18912 & 11141     & 1100822 & 22478 & 76064     & 0 & 281 & 108 \\
            & & &                               & 62.93\% & 37.07\% & 91.78\% & 1.87\% & 6.34\%   & FL & & \\ \cline{2-12}
            & Chordata & 30761 & 91785          & 10451 & 18        & 91759 & 0 & 26              & 0 & 12 & 1 \\
            & & &                               & 99.83\% & 0.49\%  & 99.97\% & 0\% & 0.03\%      & FL & & \\ \cline{2-12}
            & Nematoda & 3437 & 30127           & 21 & 3289         & 1746 & 1196 & 27185         & 1 & 2 & 11 \\
            & & &                               & 0.63\% & 99.37\%  & 5.79\% & 3.97\% & 90.23\%   & P & & \\ \cline{2-12}
            & Platyhel- & 4459 & 22683       & 7 & 7086          & 175 & 0 & 22508             & 0 & 0 & 5 \\
            & minthes & &                          & 0.1\% & 99.9\%    & 0.77\% & 0\% & 99.23\%      & P & & \\
            \hline \hline 
            \parbox[c]{1mm}{\multirow{2}{*}{\rot{\bfseries Class~}}}
            & Insecta & 91256 & 989572          & 17841 & 10734     & 1022747 & 976 & 57105       & 0 & 245 & 77 \\
            & & &                               & 62.44\% & 37.56\% & 94.63\% & 0.09\% & 5.28\%   & FL & & \\ 
            \hline
          \end{tabular} 
        \end{center}
        \caption{Some selected phyla (subtrees), only leaf nodes, are examined for the evolution 
          of the amount and percentage of given data to predicted data.}
          % \caption{All kingdoms (subtrees), only leaf nodes, are examined for the evolution of the 
          % amount and percentage of given data to predicted data.}
          % \caption{Origins and losses}
        \label{table:phylum leaf nodes states} 
      \end{table}

      \begin{table}[H]
        \centering
        \begin{tabular}{|c|l|r|r|r|r|}
        \hline
        & \multicolumn{1}{c|}{Text} & \multicolumn{1}{c|}{Text} & \multicolumn{1}{c|}{Text} & \multicolumn{1}{c|}{Text} & \multicolumn{1}{c|}{text}\\
        \hline
        \parbox[t]{2mm}{\multirow{3}{*}{\rotatebox[origin=c]{90}{rota}}} & text &&&&\\
        & text &&&&\\
        & text &&&&\\
        \hline
        \end{tabular}
        \end{table}

      In contrast to the other phyla examined, the phylogeny in the Chordata is more pronounced (less 
        multifurcation) (see Table \ref{table:percentage loss information subtrees}). This results in 
        less variance of errors. What is reflected in the results from the table 
        \ref{table:phylum leaf nodes states}. There are 18 parasites as input data and only 8 more are 
        predicted. The Chordata mostly consist of free-living species, so this seems believable. We 
        started with 99.83\%  species and predict 99.97\% species as free-living (including already 
        known nodes).

      \begin{figure}
        \centering
        \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.85\textwidth]{Figures/ChordataParasites.png}
        \caption{Parasites of Phylum: Chordata. \\
          All parasite data of the chordata are mapped into a rough taxonomy (phylum, class, order, 
            family) in order to understand their distribution and affiliation. \\
          The internal nodes are the wanted taxa from OTL (with the addition of \# input parasites to 
            $\rightarrow$ \# predicted parasites). \\
          The leaf nodes are the input parasites (green) and the predicted parasites (white $\rightarrow$ 
            green).}
        \label{fig:ChordataParasites}
      \end{figure}

      We mapped the few parasitic species in a \colorbox{red}{rough/simple} taxonomy (see Figure \ref{fig:ChordataParasites}): \\
      Known parasitic birds belong to the order Sauria. Here we know from Rothschild  here there are 
        breeding parasites, like the cuckoo and clepto-parasites as the skuas \cite{Rothschild1957}. We
        got 6 input parasites from GloBI and there are no predictions: A woodpecker - 
        \textit{Sphyrapicus varius} and a duck - \textit{Aix sponsa}, a cow bird - \textit{Molothrus ate} 
        known as broodparasite and some others. \\
      An example of the amplification of mistakes here are the carp. There is a paper from which GloBI 
        concludes: Grass carp (\textit{Ctenopharyngodon idella)} has Pathogen common carp 
        (\textit{Cyprinus carpio})\footnote{
          \hyperlink{https://www.globalbioticinteractions.org/?interactionType=hasParasite&targetTaxon=Cyprinus\%20carpio}
          {https://www.globalbioticinteractions.org/?interactionType=hasParasite\&targetTaxon=Cyprinus\%20carpio}
        }. Since there is hardly any information about free-living, it follows that all siblings are 
        also predicted to be parasitic. \\
        % Other birds: Columba livia, Nestor meridionalis, Sphyrapicus varius \\
        % Snake: Storeria dekayi (probably caused by wrong input data.)
      % \begin{itemize}
      %   \item Order Acanthomorphata(3):, one slime fish (family Blenniidae) and two others without family
      %   \item Order Anura (1): a toad (Family Bufonidae) - there exists some parasitic species in here, 
      %     they eat fin or pieces of larger fish.
      %   \item Order Siluriformes (1): a catfish (Family Loricariidae) - there exists some parasitic 
      %     species in here.
      % \end{itemize}

      The Apicomplexa are a parasitic phylum. We found only one input organism: Stemonitis fusca as a 
        free-living species. In GloBI it is listed as being parasitized by Nectria candicans and Nectriopsis 
        sporangiicola\footnote{
          \hyperlink{https://www.globalbioticinteractions.org/?interactionType=parasiteOf&targetTaxon=Stemonitis\%20fusca}
          {https://www.globalbioticinteractions.org/?interactionType=parasiteOf\&targetTaxon=Stemonitis\%20fusca}
        }. The algorithm has not predicted new free-livings. \\

      Most species of Platyhelminthes (flatworms) are parasites, although there are also free-living, 
        predatory feeding species. These are summarized in the Turbellaria, while the parasites are 
        divided into three other classes \cite{Ax1961}. This also corresponds to our observations. There 
        is one class (Rhabditophora) that contains all but one single exception of free-living species 
        of this phylum, which includes the Turbellaria. \\
      It should be noted, however, that this classification is outdated, as it has been proven that 
        the Turbellaria are not monophyletic. But we will not go into that here. \\
      For the Platyhelminthes we have more state information for the leaf nodes compared to the other 
        considered subtrees, see Table \ref{table:percentage loss information subtrees}. We start with 
        0.1\% free-livings and predicted 0.77\% as free-living species. \\

      With the Nematoda it looks more complicated. Large parts of Nematoda are free-living, but we 
        found only 5.32\% of them. Blaxter et al. estimates the order of 25,000 parasites in the 
        Nematoda \cite{Blaxter2015} and speaks of at least seven independently arosed parasitism 
        \cite{Blaxter1998}. In a recent article Blaxter identifies 18 origins \cite{Blaxter2015} in 
        Nematoda. \\
      The problem at this point, however, is: Hallan speaks of the fact that only 23,000 species 
        are described by the Nematoda but is it assumed that one million or more species are in this 
        phylum.\footnote{J. Hallan, unpublished; 
          \hyperlink{https://insects.tamu.edu/research/collection/hallan/}
          {https://insects.tamu.edu/research/collection/hallan/}
        }. \todo{Link ist nicht erreichbar!} The parasites have been much more studied and thus we start 
        with only 0.63\% (table \ref{table:phylum leaf nodes states}) free-living species. Against such 
        a shifted data situation, the algorithm is almost powerless to make correct predictions. And yet 
        the percentage has increased to 5.34\%. \\

      % \anmerkungstext{Das sind schonmal vier große Kontraste, wenn dann noch Zeit bleibt, die 
      %   schwirigen... Arthropoden, Fungi, Pflanzen... (Emanuel)} \\

      \todotext{Im Folgenden folgen 3 weitere ähnliche tabellen. Einmal eine ähnliche Tabelle inklusive 
        interner Knoten \ref{table:Phylum internal nodes}, zweimal die Übersicht über die Kingdoms 
        Blattknoten \ref{table:Kingdom leaf nodes} bzw interne Knoten \ref{table:Kingdom internal nodes}. Welche 
        nehmen wir? Rest Appendix oder ganz raus?}

      %---------------------------------------------------------------------------------------------------
      \subsubsection{Origins and Losses} \label{sec:results - origins and losses}

        Weinstein and Kuris have been searching for origins of parasitism in Animalia \cite{Weinstein2016}. 
          They identified 223 parasitic origins: 223 in Metazoa $\supset$ 143 in Arthropoda $\supset$ 87 
          in Insecta. \\
        This has led us to count the origins and losses of parasitism in our investigation as well. \\
        We count only one origin / loss in a parent node with different children's nodes. \\
        Here we have encountered a problem: The Castor algorithm gives us probabilities for states. That 
          means there are also nodes with state like 0.3 or 0.5. The question is, how to count these state 
          changes. We decided to round the values.
        % \begin{lstlisting}[gobble=6]
        %   possible states: 0, 0.333, 0.4, 0.5, 0.667, 0.75, 1
        %   rounded to:    0  0      0    0    1      1     1
        % \end{lstlisting}

        In Table \ref{table:origins and losses} we can see, that we found some more origins than Weinstein 
          and on top of that some losses. \\

        When more closely regarding the same phyla as in the section before: Chordata, Nematoda, 
          Platyhelminthes and Apicomplexa. \\
        Chordata are full of free-living species and so we see only a few origins of parasitism. The root
        and mostly all species are predicted as free-living. \\
        In Apicomplexa and the Platyhelminthes also look credible. Our algorithm gives us only one loss of
          parasitism in Apicomplexa and five in the Platyhelminthes. They are both from the root over
          mostly all species predicted as parasites. \\
        Nematoda is again full of problems. The root node is predicted as a parasite and so we have more 
          losses of parasitism for the less information of free-living species in this phylum. The rest 
          is parasitic \\
        As we have already mentioned Blaxter et al. found at least seven origins of parasitism 
          \cite{Blaxter1998}. If we assume that the root node of Nematoda is free-living, then some 
          losses would have to turn around and become Origins. So it could be that we end up in a similar 
          size as Blaxter.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Leave-100-out cross-validation} \label{sec:results - cross-validation}
      For a further validation of our results, we carried out a leave-100-out cross-validation. \\
      In order to achieve about 15\% of the 60,871 input node states with a validation, 
        we randomly left out 100 states 100 times. To make smaller amounts up to leave-one-out had too 
        much computational effort. \todo{warum 15\%?}\\

      Of these 10,000 nodes, 9,238 are unique. From the unique nodes, we predicted approximately 
        98.17\% correct and thus 1.82\% wrong, with duplicate draws always having the same prediction 
        (data not shown). \\

      We have again dealt with the question of how this data is best modeled. The influence of the taxa 
        (kingdom, phylum, class) and the depth of the leaf nodes is re-modeled and the BICs compared 
        (table \ref{table:BIC cross-validation}). Lower taxa than classes (e.g. order) are 
        computationally too expensive to calculate. \\

      \todo{table for these numbers?}
      % 100/(25962+34860)*10000 = 16.44
      % 100/(25962+34860)*9238 = 15.19
      % 10000 nodes 9238 unique
      % 9060 nodes $\approx 98.17\%$ correctly predicted
      % 169 ($\approx 1.82\%$) wrongly predicted

      \begin{table}[h]
        \begin{center}
          \begin{tabular}{ |l|r|r|r| }
            \hline
            \bfseries Model / Taxa & \bfseries Kingdom & \bfseries Phylum & \bfseries Class \\% & Order \\
            \hline \hline
            correctly predicted $\sim$ taxa & 117936 & \cellcolor{green!40}112242 & \cellcolor{green!50}111733 \\% & XXXXX \\
            \hline
            correctly predicted $\sim$ taxa + depth & 117776 & \cellcolor{green!50}111304 & \cellcolor{green!50}111273 \\% & XXXXX \\
            \hline
            correctly predicted $\sim$ taxa * depth & 117709 & \cellcolor{green!50}111262 & \cellcolor{green!30}113135 \\% & XXXXX \\
            \hline
          \end{tabular} 
        \end{center}
        \caption{BIC (Bayesian information criterion) values of cross-validation prediction models. \\
          These models are created with the R function \textit{glm()} and compared with the 
            \textit{BIC()} function. This results in the listed BIC values, presented in a heatmap colorization.}
        \label{table:BIC cross-validation} 
      \end{table}

      The BIC values this time did not prove that the finest model is the best. Of our calculated models, 
       \textit{correctly predicted $\sim$ phylum * depth} has the smallest value. \\

      We examined the influence of the omitted data on the prediction. On average, about twice as many 
        leaf nodes are predicted differently. \color{red}The variance is very high. This means that 
        we have a high degree of dispersion and thus a stochastic situation exists. It also describes 
        the width of the present probability function. The standard deviation is about five times as 
        high as the number of omitted nodes, so the variability is quintupled. \color{black} Table 
        \ref{table:statistics cross-validation} shows these results. \\
      \begin{table}[h!]
        \begin{center}
          \begin{tabular}{ |cl||r|r|r|r|r| }
            \hline
            & & \bfseries min & \bfseries max & \bfseries mean & \bfseries variance ($\sigma^2$) & $\sigma$ \\
            \hline \hline
            \multirow{3}{*}{\bfseries distance} & \bfseries all     & 0 & 3587.70 & 224.96 & 313650.61 & 560.05 \\
            & \bfseries leaf nodes                                  & 0 & 3021.12 & 208.69 & 248103.38 & 498.10 \\
            & \bfseries internal nodes                              & 0 & 566.58 & 16.28 & 4927.95 & 70.20 \\ \hline
            \multirow{3}{*}{\bfseries lost} & \bfseries all states  & 100 & 100 & 100 & 0 & 0 \\
            & \bfseries FL states                                   & 44 & 66 & 57.25 & 19.50 & 4.42 \\
            & \bfseries P states                                    & 34 & 56 & 42.75 & 19.50 & 4.42 \\
            \hline
          \end{tabular}
        \end{center}
        \caption{Statistics about the leave-100-out cross-validation \\
          The distance between original and new states is calculated using the Euclidean metric. This 
            is summed over all states / all leaf node states and all internal node states. \\
          The lower half of the table describes the distribution of the 'lost' states between parasites 
            (P) and free-livings (FL).}
        \label{table:statistics cross-validation}
      \end{table}

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Discussion}

  \todo{Aus den models kann man auch schließen: highly variable throughout our tree}

  \todo{noch irgendwo einfügen:} \\
  The number of nodes in the tree corresponding to a taxonomic rank is tabulated. \\
  For the classical taxonomic ranks phylum, class, order, family, genus xyz is found (appendix 5.2).
  % A closer look is taken at some of the features of the synthesis tree on the one hand the 
  % distribution of the taxa and on the other the distribution of the nodes on the taxa. Since this 
  % is \colorbox{red}{not directly relevant} for this study, this section has been moved to the appendix 
  % \ref{sec:otl analysis}.

  This chapter deals with the question of how trustworthy our result of the ancestral state 
    reconstruction of the Eukaryota tree is. And further on how well our simulation can simulate the 
    real problem and thus make statements about the predictive power of the applied Sankoff algorithm.

  We have pursued this question in various investigations and yet, of course, further possibilities 
    for improvement remain. Despite these possibilities for improvement, our reconstruction gives a 
    first good assessment of the whole tree.

  \todo{Why is the study important?} \\
  \todo{How does this study relate to previous studies?}

  We evaluated our ancestral state reconstruction and the relational state prediction using three 
    methods: a simulation, a selective analysis from a biological perspective and a leave-100-out 
    cross-validation.

  Each theme has its own limitations. \\
  There are various expansion options and points of criticism for the simulation, which in 
    particular revolve around the question of how well we can simulate our real data. We dedicate 
    this question to the subsection \ref{sec:discussion - simulation}. \\
  Above all, the 'biological view' gives a first impression of the credibility of the results. But 
    he can not make any statistical statements. Here the interested researcher may have to analyze 
    the subtrees of relevant to him for himself. \\
   Of course, cross-validation can be expanded. With more computing power, a leave-1-out cross 
    validation could also be performed. However, our result on this point yields a strong result 
    with 98.17\% correct prediction that this does not seem necessary. \\
  One last limitation also exists in the data situation itself. We have pointed out in various 
    sections that we have a strong multifurcation and only a few data about leaf node states. In 
    figure \ref{fig:influence of unknown data} you can see how the predictive power can improve if 
    we have more data available. We discuss this point in section 
    \ref{sec:discussion - data situation}.

  % Wieviele Origins haben wir gefunden, was bedeutet diese Zahl? \\

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Data situation} \label{sec:discussion - data situation}
    The used eukaryotic synthesis tree from OTL \cite{Hinchliff2015} has 241,974 internal nodes and 
      293,463 leaf nodes. In addition, we could specify 25,962 parasitic and 34,860 free-living 
      species from GloBI \cite{Poelen2014}. \\
    This gives us a high number of missing internal nodes (high multifurcation) and a low number of 
      node states of the leaf nodes. In table \ref{table:percentage loss information subtrees} we have 
      written down this percentage including some subtrees. \\

    The investigation of the taxonomy revealed that the OTL tree has three kingdoms: Chloroplastida, 
      Metazoa, Fungi, 53 phyla, 195 classes and 924 orders. \\
    Since the analysis of the tree is not part of this work, it should be mentioned here that, 
      according to recent findings, this is not complete and we lack some taxa in every rank. For 
      example, Cavalier-Smith says that one distinguishes between seven and nine kingdoms 
      \cite{CavalierSmith1981}. \\
    In \colorbox{red}{section/page} \pageref{subsec:listPhyla} of the appendix is a list of all phyla. \\

    Also, the global database is not infallible. We found out (section \colorbox{red}{x}) that there is some 
      misinformation. \\
    We are able to report some of these directly to Globi.

    We found also 57,352 (not distinct!) source species and 809,993 (not distinct!) target species 
      without OTT indentifyers. Since we currently use only OTT indentifyers, we could not use this information. \\
    At this point there is thus the possibility to use more of the existing data, if one performs a 
      matching with the other indentifyers.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Simulation} \label{sec:discussion - simulation}

\todo{einbauen....}
  As mentioned above, in an ancestral state reconstruction, transitions are minimized, with 
  transition probabilities playing a role. \\
\colorbox{red}{This subsection deals with} these transition probabilities from free-living (hereinafter / as a 
  formula FL) to parasitic (hereinafter P) and vice versa: $\mathcal{P}(FL \rightarrow P)$, 
  $\mathcal{P}(P \rightarrow FL)$. \\
Different parasite types have different transition probabilities. It is very difficult to make a 
  statement about these probabilities. It is generally assumed in this work that there are 40\% 
  parasites and 60\% free-livings which is based on the estimates by Windsor \cite{Windsor1998} 
  and $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$, as a result of the 
  diversity of parasites and the lack of general determinations for this. These parameters are 
  debated in subsection \ref{sec:discussion - transition probabilities} of the discussion. \\

  However, the used castor package \cite{Louca2017} offers the possibility to enter different 
  transition probabilities.

  Wieviele Übergänge gab es dabei? Hast du die gleich (ähnlich groß) gehalten?
  Der Anteil von Parasiten und Freilebenden ist ja nicht gleich der Transition probability. Das sind zwei verschiedene Dinge...auch wenn die Transition probability den Anteil ergibt. 
  neben dem Anteil der P/FL ist auch die Anzalhl der Übergänge zu berücksichtigen!
  50/50 kann man in einem Baum mit 1000 nodes mit einem einzigen Übergan erreichen, der die eine Häfte des Baumes ändert.
  50/50 kann man aber genauso gut mit einem Übergan auf JEDER edge erreichen. Dann würde sich immer ein P mit einem FL knoten abwechselnn....

... für die Diskussion!

\todo{-----}

    The aim of the simulation is to test the influence of various unknown or unsafe parameters in 
      order to test the credibility of the prediction. \\
    Some of these influences could not be tested: Firstly, the distribution of parasites and 
      free-livings in the tree produces various parameters that we could not test all (see subsection 
      \ref{sec:discussion - transition probabilities}) and, secondly, we discuss the simulation of a 
      phylogenetic tree on which the evaluation based on certain parts (see subsection 
      \ref{sec:discussion - simulate phylogeny}).

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Transition probabilities} \label{sec:discussion - transition probabilities}
      We assume that different parasite types have different transition probabilities. Establishing a 
        common distribution across the eukaryotic tree is therefore difficult. \\
      Based on the estimates of Windsor \cite{Windsor1998}, we have assumed a distribution of 40\% 
        parasites to 60\% free livings in this work. As a result of the diversity of parasites and the 
        lack of generalizations, we have generally stated that 
        $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$. But it is also reasonable to 
        assume that in general $\mathcal{P}(FL \rightarrow P) > \mathcal{P}(P \rightarrow FL)$, because 
        a reverse mutation is usually less likely. However, one would have to determine how much this 
        difference is and thus discuss another parameter. \\

      In the simulation, we tested different beta distributions with different thresholds and found that 
        this combination has a considerable influence on the predictive power of the algorithms. \\
      At this point it would be possible to test other distributions with equal threshold values or 
        different threshold values with equal distributions. \\
      However, if you choose a strong free-living distribution and balance it with the threshold to 
        achieve a 40:60 parasite free-living distribution, for example, this could be very 
        computationally expensive. \\

      With these further simulations you could find out if the issue of distributions plays a big role. 
        Conversely, one could estimate possible distributions based on the data location in the tree. 
        However, this is likely to be very difficult given the poor data. For this reason, we have 
        decided in this work to accept very general values and not to speculate much. \\

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Simulation of a phylogeny} \label{sec:discussion - simulate phylogeny}
      In our simulation, we start with the simulation of a phylogeny and then depict our data situation. \\
      For this first step we use the \textit{randomized()} function of the phylo package from biopython
        \cite{Cock2009}. The question is how close is such a randomized tree to a phylogeny. \\

      The problems that can arise are that different species develop at different rates. This means that 
        there could be sections in the tree with many branches (for example, subtrees of unicellular 
        species), ie with high depths and opposite sections (of very complex species, for example). It 
        can thus be assumed that the tree is not balanced. \\
      For more precise statements, one would have to take deeper undercurrents, which was not possible 
        in the context of this work.

      \todotext{It should be noted that we have optimized the distributions to the proportions and thus the number of state changes will be relatively high. With a smaller number of state changes, the predictive power of both algorithms would be even better.}
      % transition probability

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Conclusion}

    From the simulation, we can conclude that we predict correctly about 60\% of the nodes in the 
      present data situation (see section \ref{sec:results - simulation}).
    The leave-100-out cross-validation even showed that we predict the omitted nodes to be 98.17\%  
      percent correct (see section \ref{sec:results - cross-validation}). \\
    This allows for the assumption that the data is grouped and not uniformly distributed and thus 
      smaller subtrees are present in which data are to be found in the simulation with smaller 
      multifurcation and smaller value for unknown nodes. This is also confirmed by our biological 
      analysis of the Chordata subtree (see subsection \ref{sec:results - biological view}). \\
    However, this means that the ancestral states' data in the direction of root node are probably 
      particularly unbelievable. Which makes the localization of Origins direction root node 
      difficult. The question remains, how much this affects the estimation of the number of origins. 
      Our comparison with the paper from Weinstein and Kuris \cite{Weinstein2016} (see subsection 
      \ref{sec:results - origins and losses}), however, leaves us with optimism.

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\bibliography{bibliographie}

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
\chapter{Appendix}
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Methods overview}
    % \begin{figure}[h!]
    %   \centering
    %   \includegraphics[angle=90, width=0.4\textwidth]{Figures/Workflow.png}
    %   \caption{Big overview of the whole Workflow}
    %   \label{fig:BigWorkflow}
    % \end{figure}
    \begin{figure}[h!]
      \centering
      \includegraphics[width=\textwidth]{Figures/Workflow.png}
      \caption{Big overview of the whole Workflow}
      \label{fig:BigWorkflow}
    \end{figure}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{OTL analysis}\label{sec:otl analysis}

    %---------------------------------------------------------------------------------------------------
    \subsection{List of all phyla}\label{subsec:listPhyla}

    Phyla (53): \\
    Acanthocephala, Amoebozoa, Apicomplexa, Arthropoda, Ascomycota, Bacillariophyta, Basidiomycota, 
      Brachiopoda, Bryozoa, Chaetognatha, Chlorophyta, Chordata, Chromerida, Chytridiomycota, 
      Ciliophora, Cnidaria, Colponemidia, Ctenophora, Cycliophora, Echinodermata, Entoprocta, 
      Entorrhizomycota, Euglenida, Foraminifera, Gastrotricha, Glomeromycota, Gnathostomulida, 
      Haplosporida, Haptophyta, Hemichordata, Kinorhyncha, Loricifera, Microsporidia, Mollusca, 
      Myzostomida, Nematoda, Nematomorpha, Nemertea, Onychophora, Orthonectida, Phaeophyceae, 
      Picozoa, Placozoa, Platyhelminthes, Porifera, Priapulida, Rhodophyta, Rhombozoa, Rotifera, 
      Streptophyta, Tardigrada, Xanthophyceae \\
    Wobei von Streptophyta -> Anthocerotophyta, Marchantiophyta, Bryophyta, Tracheophyta als
      Phylum im Phylum gefunden und nicht einbezogen wurden und Magnoliophyta als Phylum in 
      Tracheophyta ebenfalls nicht. \\

    %---------------------------------------------------------------------------------------------------
    \subsubsection{Distribution of Taxa}
    - In the tree we can distinguish 28 different Taxa with the OTL taxonomic tree. \\
    - The most of them are hardly represented. The major taxonomic groups are: ... \\
    - Here \colorbox{red}{you} can see some characteristics of the Multifurcation of the tree. \\
    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.9\textwidth]{Figures/TaxaTable2.JPG}
      \caption{Distribution of Nodes in Rank-Cathegories}
      \label{fig:taxaTable2}
    \end{figure}
    In a phylogeny, the taxonomic division of the tree is far too coarse, meaning that there should 
      be more subtaxa or 'unranked' nodes. But the closer we get to the root, the more the pure
      taxonomic tree is reflected. If the tree are binary, the taxa would have to double. But the 
      multipliers for some are much bigger and for others much smaller, which \colorbox{red}{you} can see in in figure 
      \ref{fig:taxaTable2}. \\
      ... (see Table \ref{table:...})

    \begin{table}[h!]
      \begin{center}
        % \begin{tabular}{ |>{\rowmac}l|>{\rowmac}r|<{\clearrow} }
        \begin{tabular}{ |l|r||r|r|r| }
          \hline
          \bfseries Taxa & \bfseries Number of Nodes & \bfseries Internal Nodes & \bfseries Leaf Nodes & \bfseries Terminal Leaf Nodes \\
          \hline \hline
          \setrow{\bfseries}domain & 1 & 1 & &  \\ \hline
          \setrow{\bfseries}kingdom & 3 & 3 & &  \\
          subkingdom & 2 & 2 & & \\
          infrakingdom & 5 & 5 & & \\
          superphylum & 2 & 2 & & \\ \hline
          \setrow{\bfseries}phylum & 57 & 54 & 3 & \\
          subphylum & 22 & 20 & 2 & \\
          infraphylum & 2 & 2 & & \\
          superclass & 8 & 8 & & \\ \hline
          \setrow{\bfseries}class & 209 & 175 & 34 & \\
          subclass & 120 & 106 & 14 & \\
          infraclass & 14 & 11 & 3 & \\
          superorder & 40 & 37 & 3 & \\ \hline
          \setrow{\bfseries}order & 1014 & 772 & 242 & \\
          suborder & 285 & 222 & 63 & \\
          infraorder & 95 & 78 & 17 & \\
          parvorder & 14 & 12 & 2 & \\
          superfamily & 829 & 626 & 203 & \\ \hline
          \setrow{\bfseries}family & 8449 & 6143 & 2306 & \\
          subfamily & 3090 & 2250 & 840 & \\
          supertribe & 1 & 0 & 1 & \\
          tribe & 1764 & 1285 & 479 & \\
          subtribe & 435 & 359 & 77 & \\ \hline
          \setrow{\bfseries}genus & 164656 & 105452 & 59204 & \\
          subgenus & 1266 & 869 & 397 & \\
          section & 3 & 2 & 1 & \\
          subsection & 1 & 1 & 0 & \\
          species group & 365 & 288 & 77 & \\
          species subgroup & 123 & 100 & 23 & \\ \hline
          \setrow{\bfseries}species & 2247251 & 7423 & 2239828 & 2228993 \\
          subspecies & 22437 & 75 & 22362 & 22239 \\
          variety & 1755 & 7 & 1748 & 1726 \\
          varietas & 5970 & 19 & 5951 & 5909 \\
          forma & 1181 & & 1181 & 1181 \\
          \hline \hline
          no rank & 954 & 719 & 235 & 7 \\
          no rank - terminal & 37452 & & 37452 & 37452 \\
          (no entry) & 40099 & 40099 & & \\
          \hline  
        \end{tabular}
        \caption{\todo{...}}
        \label{table:...} 
      \end{center}  
    \end{table}
    extended leaf nodes (real leaf nodes)

  %---------------------------------------------------------------------------------------------------
  \subsubsection{Distribution of data in the taxa}
    Mithilfe des taxonomischen Baums von OTL haben wir die Knoten ihren Kingdoms, Phyla und Classes 
      zugeteilt (see Table \ref{table:...}). \\

    \begin{table}[h]
      \begin{center}
        \begin{tabular}{ |l|r|l|r|r| }
          \hline
          \bfseries Kingdom (3) & \bfseries Number of Nodes & \bfseries Phylum (25) & \bfseries Number of Nodes & \bfseries max max height \\
          \hline \hline
          Metazoa & 1 465 207 & Arthropoda & 1 170 539 & 54 \\
          & & Chordata & 106 650 & 74 \\
          & & Mollusca & 80 022 & 22 \\
          & & Platyhelminthes & 27 141 & 9 \\
          & & Nematoda & 24 564 & 23 \\
          & & Cnidaria & 14 878 & 36 \\
          & & Porifera & 11 737 & 26 \\
          & & Echinodermata & 10 654 & 14 \\
          & & Bryozoa & 8 631 & 11 \\
          & & Rotifera & 3 093 & 7 \\
          & & Nemertea & 1 793 & 7 \\
          & & Tardigrada & 1 654 & 7 \\
          & & Acanthocephala & 1 596 & 6 \\
          & & Brachiopoda & 1 055 & 9 \\
          & & Nematomorpha & 633 & 7 \\
          & & Chaetognatha & 360 & 7 \\
          & & Hemichordata & 196 & 5 \\ 
          & & Cycliophora & 11 & 5 \\ 
          % 1465207-(1170539+106650+80022+27141+24564+14878+11737+10654+8631+3093+1793+1654+1596+1055+633+360+196+11) = 0          \hline
          Fungi & 254 871 & Ascomycota & 157 045 & 19 \\ 
          & & Basidiomycota & 92 931 & 18 \\
          & & Microsporidia & 1 949 & 6 \\ 
          & & Glomeromycota & 1 490 & 6 \\
          & & Chytridiomycota & 1 456 & 6 \\
          % 254871-157045-92931-1949-1490-1456 = 0
          \hline
          Chloroplastida & 121 239 & Streptophyta & 120 731 & 49 \\
          & & Chlorophyta & 508 & 6 \\
          % 121239-120731-508 = 0
          \hline  
        \end{tabular}
        \caption{\todo{...}}
        \label{table:...} 
      \end{center}  
    \end{table}

  %---------------------------------------------------------------------------------------------------
  \section{Results}

    \begin{table}[h!]
      \begin{center}
        \begin{tabular}{ |l|r||r|r||r|r|r|r|r|r| }
          \hline
          & & \multicolumn{2}{c||}{\bfseries original states} & \multicolumn{6}{c|}{\bfseries final states} \\
          \bfseries Phylum & \bfseries \# nodes & \bfseries FL & \bfseries P
            & \bfseries 0 (FL) & \bfseries 0.4 & \bfseries 0.5 & \bfseries 0.67 & \bfseries 0.75 & \bfseries 1 (P) \\
          \hline \hline
          Chordata & 122546 & 10451 & 18 
            & 122473 & 0 & 0 & 0 & 0 & 73 \\
          Nematoda & 33564 & 21 & 3289 
            & 846 & 0 & 1133 & 0 & 0 & 31585 \\
          Platyhelminthes & 27142 & 7 & 7086 
            & 1010 & 0 & 175 & 0 & 0 & 25957 \\
          Apicomplexa & 2102 & 1 & 255 
            & 1 & 0 & 0 & 0 & 0 & 2101 \\
          % \hline
          % Arthropoda & 1319460 & 18912 & 11141 
          %   & 1207204 & 1313 & 25499 & 4852 & 1957 & 78635 \\
          \hline
        \end{tabular}
      \end{center}
      \caption{Some selected phyla (subtrees), including internal nodes, are examined for the evolution 
      of the amount and percentage of given data to predicted data.}
      \label{table:Phylum internal nodes}
    \end{table}

    \begin{table}[h!]
      \begin{center}
        \hspace*{-2cm}\begin{tabular}{ |l|r||r|r||r|r|r|r|r|r|r|r| }
          \hline
          & & \multicolumn{2}{c||}{\bfseries original states} & \multicolumn{8}{c|}{\bfseries final states} \\
          \bfseries Kingdom & \bfseries \# nodes & \bfseries FL & \bfseries P
            & \bfseries 0 (FL) & \bfseries 0.25 & \bfseries 0.33 & \bfseries 0.4 & \bfseries 0.5 & \bfseries 0.67 & \bfseries 0.75 & \bfseries 1 (P) \\
          \hline \hline
          none & 84456 & 45 & 529 
            & 15035 & 243 & 25910 & 0 & 8764 & 6183 & 0 & 28140 \\
          Fungi & 324105 & 577 & 2983
            & 39088 & 0 & 0 & 0 & 5858 & 0 & 0 & 274803 \\
          Chloroplastida & 460457 & 3519 & 77
            & 454211 & 0 & 0 & 0 & 4688 & 0 & 0 & 1558 \\
          Metazoa & 1670956 & 30758 & 22373
            & 1485749 & 0 & 0 & 1313 & 29002 & 5102 & 1957 & 147833 \\
          \hline  
        \end{tabular}
      \end{center}
      \caption{All kingdoms (subtrees), inkluding internal nodes, are examined for the evolution of the 
      amount and percentage of given data to predicted data.}
      \label{table:Kingdom internal nodes}
    \end{table}

\end{document}